---
title: "Peer Assessment 1"
author: "jagdeep s sihota"
date: "August 11, 2014"
output: html_document
---

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

Data

The data for this assignment can be downloaded from the course web site:
Dataset: Activity monitoring data [52K]


Loading and preprocessing the data


```{r}
if (!file.exists("activity.csv")) {
    unzip("activity.zip")
}
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", 
    "numeric"))
activity$date <- as.Date(activity$date, "%Y-%m-%d")
totalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)

hist(totalSteps$steps, main = "Histogram of the total number of steps taken each day", xlab = "total number of steps taken each day", col = "green")
```


What is mean total number of steps taken per day?
And the mean and median is
```{r}
  mean(totalSteps$steps)
  median(totalSteps$steps)
```


What is the average daily activity pattern?

Make a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Now I get mean of steps and time series plot

```{r}
time_series <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
```
The plot is:
```{r}
plot(row.names(time_series), time_series, type = "l", xlab = "Time Series Plot per 5-minute interval", 
    ylab = "Average across all Days", main = "Average number of steps taken", 
    col = "green")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- which.max(time_series)
names(max_interval)
```
Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
activity_NA <- sum(is.na(activity))
activity_NA
```

missing values are replaced with average of steps per interval
```{r}
meanInterval <- aggregate(activity$steps, by = list(activity$interval), mean, na.rm = TRUE, 
    na.action = NULL)
names(meanInterval)[1] <- "interval"
names(meanInterval)[2] <- "mean.steps"
naRepacedData <- merge(activity, meanInterval, by = "interval", sort = FALSE)  
naRepacedData <- naRepacedData[with(naRepacedData, order(date, interval)), ]  
naRepacedData$steps[is.na(naRepacedData$steps)] <- naRepacedData$mean.steps[is.na(naRepacedData$steps)]
nAReplacedData <- naRepacedData[, c(2, 3, 1)]
```
plot
```{r}
activityDataWithNAReplaced <- aggregate(nAReplacedData$steps, by = list(nAReplacedData$date), sum)
names(activityDataWithNAReplaced)[1] <- "day"
names(activityDataWithNAReplaced)[2] <- "steps"
hist(activityDataWithNAReplaced$steps, main = "Histogram of the total number of steps taken each day (NA replaced)", xlab = "total number of steps taken each day", col = "green")
```
Get mean
```{r}
mean(activityDataWithNAReplaced$steps, na.rm = TRUE)
```
Get median
```{r}
median(activityDataWithNAReplaced$steps, na.rm = TRUE)
```

The Mean is equal to the estimates from the first part of the assignment.

The Median is slightly lower when compared to the first part of the assignment.

The histogram shows a similar shape as before with overall higher frequencies due to the NA being replaced in the new histogram. See also this side by side plot:
```{r}
activityWithNa <- aggregate(activity$steps, by = list(activity$date), sum)
names(activityWithNa)[1] <- "day"
names(activityWithNa)[2] <- "steps"
par(mfrow = c(1, 2))

hist(activityWithNa$steps, main = "(with NA)", xlab = "total number of steps taken each day")

hist(activityDataWithNAReplaced$steps, main = "(NA replaced)", xlab = "total number of steps taken each day")
```

Are there differences in activity patterns between weekdays and weekends?
```{r}
day <- weekdays(activity$date)
daylevel <- vector()
for (i in 1:nrow(activity)) {
    if (day[i] == "Saturday") {
        daylevel[i] <- "Weekend"
    } else if (day[i] == "Sunday") {
        daylevel[i] <- "Weekend"
    } else {
        daylevel[i] <- "Weekday"
    }
}
activity$daylevel <- daylevel
activity$daylevel <- factor(activity$daylevel)

stepsByDay <- aggregate(steps ~ interval + daylevel, data = activity, mean)
names(stepsByDay) <- c("interval", "daylevel", "steps")
```
Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r}
library(lattice)
xyplot(steps ~ interval | daylevel, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")
```